---
title: "Assignment 14 Using XGBoost Model"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document: default
---

```{r, echo=TRUE, message=FALSE, warning=FALSE}
# Load necessary libraries
library(dplyr)
library(ggplot2)
library(xgboost)
library(caret)
library(data.table)
library(rvest)
```

Loading the given dataset as mydata dataframe
```{r, echo=TRUE, message=FALSE, warning=FALSE}
# Load the given dataset
mydata <- fread('/Users/utsingh/Downloads/flinfo2022.csv')
# Display header and first few rows of mydata
head(mydata)
```

Data cleansing
```{r}
# Delete the unassigned county name
mydata <- subset(mydata, (Admin2 != "Unassigned"))
# Drop unnecessary columns
columns_to_drop <- c('V1', 'FIPS', 'Province_State', 'Country_Region', 'Last_Update', 'Latitude', 'Longitude', 'Recovered', 'Active', 'Combined_Key')
mydata <- mydata %>% select(-all_of(columns_to_drop))
head(mydata)
```

Loading hospital beds data and adding that to mydata
```{r}
# Load hospital beds data (extracted from this website- https://www.flhealthcharts.gov/ChartsReports/rdPage.aspx?rdReport=NonVitalInd.DataViewer&cid=313)
hosp_beds <- fread('/Users/utsingh/Downloads/florida_hospital_beds_cleaned_2020.csv')

# Rename County column name to ensure correct merge
colnames(hosp_beds)[colnames(hosp_beds) == 'County'] <- 'Admin2'

# Merging hospital beds data with mydata
mydata <- merge(mydata, hosp_beds, by = "Admin2", all.x = TRUE)
head(mydata)
```

Calculating and saving beds per population
```{r}
# Create new feature (beds per population)
mydata <- mydata %>%
  mutate(
    beds_per_popu = ifelse(population > 0, Hospital_Beds / population, NA)  # Avoid division by zero
  )
head(mydata)
```

Loading poverty rate data and adding that to mydata
```{r}
# Load Poverty rate data (extracted from this website- https://www.flhealthcharts.gov/ChartsDashboards/rdPage.aspx?rdReport=NonVitalInd.Dataviewer&cid=294)
Poverty_Rate <- fread("/Users/utsingh/Downloads/HDPulse_data_export.csv")
# Remove the word “County” from the values in the Admin2 column
colnames(Poverty_Rate)[colnames(Poverty_Rate) == 'County'] <- 'Admin2'

# Match the content of mydata
Poverty_Rate$Admin2 <- gsub(" County", "", Poverty_Rate$Admin2)

# Left join
mydata <- left_join(mydata, Poverty_Rate, by = "Admin2")
head(mydata)
```

Loading county wise population density data and adding that to mydata
```{r}
# getting other data - land area
library(rvest)
webpage <- "http://www.usa.com/rank/florida-state--land-area--county-rank.htm"
read_html(webpage)

url <- webpage
pageobj <- read_html(url, as.data.frame=T, stringsAsFactors = TRUE)
pageobj %>%  
  html_nodes("table") %>% 
  #Indicate that this is the table to be extracted
  .[[2]] %>%
  html_table(fill=T) -> x
coltitle <- as.character(x[1,]) # read the column names from row 1
colnames(x) <- coltitle # add the column names to x
x <- x[-1,-1] # remove the first row and the 1st column of the data
landarea <- as.data.frame(x)
colnames(landarea) <- c("SQMiles", "Admin2")
# clean up the data of landarea
landarea$SQMiles <- as.numeric((gsub("[\\$, sq mi]", "", landarea$SQMiles)))
landarea$Admin2 = as.character(gsub(", FL /", "", landarea$Admin2))
landarea$Admin2 = as.character(gsub(",", "", landarea$Admin2))
landarea$Admin2 = as.character(gsub(" [0-9]", "", landarea$Admin2))
landarea$Admin2 = as.character(gsub("[0-9]", "", landarea$Admin2))
landarea$Admin2[landarea$Admin2 == "Saint Johns"] <-"St. Johns"
landarea$Admin2[landarea$Admin2 == "Saint Lucie"] <-"St. Lucie"
landarea$Admin2[landarea$Admin2 == "De Soto"] <-"DeSoto"
# add the land area to the data frame
mydata <- merge(x=mydata, y=landarea, by="Admin2", all.x = TRUE)

# create a new variable that is population density
mydata$pop_density <- mydata$population / mydata$SQMiles
head(mydata)
```

Loading median age data and adding that to mydata
```{r}
# Load median age (extracted from this website- https://www.flhealthcharts.gov/ChartsDashboards/rdPage.aspx?rdReport=NonVitalIndRateOnly.TenYrsRpt&cid=300)
med_age <- read.csv("/Users/utsingh/Downloads/Median_age_FL.csv")

# Match the content of mydata
colnames(med_age)[colnames(med_age) == 'X'] <- 'Admin2'
colnames(med_age)[colnames(med_age) == 'X2022'] <- 'median_age'

# Left join
mydata <- left_join(mydata, med_age, by = "Admin2")
head(mydata)
```
Check Summary of mydata
```{r}
summary(mydata)
```
Select columns to be used for the XGBoost model
```{r}
# Select features for the model 
X_features <- mydata %>%
  select(beds_per_popu, poverty_rate, pop_density, median_age)
y <- mydata$Case_Fatality_Ratio
```

Convert data to matrix format
```{r}
# Convert to matrix format for XGBoost
X_matrix <- as.matrix(X_features)
y_vector <- as.numeric(y)
```

Train the model
```{r}
# Train XGBoost model
xgboost_model <- xgboost(
  data = X_matrix,
  label = y_vector,
  nrounds = 30,
  params = list(
    eta = 0.08,
    max_depth = 2,
    min_child_weight = 3,
    objective = 'reg:squarederror',
    eval_metric = 'rmse'
  ),
  verbose = FALSE
)
```

Calculate R-squared value to check model accuracy and print it
```{r}
# Make predictions and calculate R²
predictions <- predict(xgboost_model, X_matrix)
SSE <- sum((y_vector - predictions) ^ 2)
SST <- sum((y_vector - mean(y_vector)) ^ 2)
R_squared <- 1 - (SSE / SST)

# Print R² value
cat("R² for the Case Fatality Rate prediction is:", R_squared, "\n")
```
Plot Residual Plot
```{r}
# Residual plot
residuals <- y_vector - predictions
ggplot(data.frame(predictions, residuals), aes(x = predictions, y = residuals)) +
  geom_point() +
  geom_hline(yintercept = 0, linetype = "dashed", color = "orange") +
  labs(title = "Residual Plot", x = "Predicted Values", y = "Residuals")
```
Plot Actual vs. Predicted plot
```{r}
# Actual vs. Predicted plot
ggplot(data.frame(actual = y_vector, predicted = predictions), aes(x = actual, y = predicted)) +
  geom_point() +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "blue") +
  labs(title = "Actual vs. Predicted", x = "Actual Values", y = "Predicted Values")
```
